---
output: html_document
editor_options: 
  chunk_output_type: console
---


# Create book infrastructure

https://bookdown.org/yihui/bookdown/get-started.html



## Change index.rmd

* Title
* Author
* Description

## Update README.md

## Update _output.yml

Changes basic features of this file as needed


## Use subdirectories

To oranize .rmd files in subdirectories change _bookdown.yml to include
```{r, eval = F}
rmd_subdir: true
```



## Keep the merge file 

For reference this can be useful

in _bookdown.yml chane
delete_merged_file: true


## Workflow

### Build book

Use "Build book" button on Build tab of Env / Hist / Connections / Build / Tutorial

### Clean up if needed

```{r, eval = F}
rmarkdown::clean_site(preview = FALSE)
```

## Set up options for entire

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.height = 3, 
                      fig.width = 6, 
                      warning = FALSE,
                      message = FALSE,
                      error = FALSE)
suppressMessages(library(ggplot2))
suppressMessages(library(ggpubr))

packageStartupMessage(library(ggplot2))
packageStartupMessage(library(ggpubr))

```


## Skip files

put underscore in front of file name, eg.

_skipme.Rmd

## Suppress warnings etc

In chunk:
message=FALSE, warning=FALSE

suppressMessages(library(foo))

## Do not stop on error

https://bookdown.org/yihui/rmarkdown-cookbook/opts-error.html

```{r, error=TRUE}
1 + "a"
```
