# Scatter plots

```{r}
library(compbio4all)
```


## Checkpoint: Plot the data

Load the data and make a plot that matches Figure 2.8 on page 26 of the Higgs and Atwood chapter 2 reading

In order to make this plot, you will have to

* Read Chapter 
* Examine Figure 2.8 carefully
* Examine Table 1 of Higgs 2009, especially the notes at the bottom of the table under the data themselves
* Use the basic R code below and change the variables being plotted
* Label the axes as in Chapte 2.
* Render your plot to RPubs, screen grab it, and submit it.
* Contact me AND your UTA if/when you run into trouble.
* The UTAs and I all have office hours on Monday

The plot() function is set up to make a blank plot.  text() then uses the same tilda notation as plot() to plot the labels.

```{r}
par(mfrow = c(1,1))
plot(polar.req  ~ freq, data = aa_dat,
     xlab = "label here",
     ylab = "label here",
     main = "title here",
     col = 0)
text(polar.req ~ freq, 
     labels = aa, 
     data = aa_dat, 
     col = 1:20)
```
