# Working outline {-}

## TODO

add to compbio4all package

* frogarms 
* sparrow telomeres
* gene size distribution


## General book conventions

* "BOOK" = major unit; each BOOK in own folder
  + Book names are generaly based on important / interesting books related to R, data science, quantitatitve biology, etc.
* "Part" = major subunit; each Part in own subfolder
* Chapter = individual lesson/exercise; separate .Rmd files

Many chapters have sections labled **NEW text** followed by one flagged **OLD text**.  The old text is from my previous book project and likely needs to be re-written and aligned with the new format.

## Book names to use / riff off of for BOOK names

* Tufte: "The Visual Display of Quantitative Information", "Beautiful evidence", "Visual Explanations", "Envisioning information"

## Outline

Not necessarily up to date with current structure of books

### BOOK 1: Introduction: Programming with data (Chambers 1998)

General info on R, bioinformatics, comp. bio., the book etc

### BOOK 2: "Biology by the numbers" (Burton)

Actual title of Burton: "Biology by numbers"

* Plotting, data summary


#### Part 1: (Time series plots, and lots of bits and pieces)

**Implemented**

* Time series plots - base R

**Not implemented**

* Cumulative number of amino acids discovered 
* Clines?
* Eagle data - 


#### Part 2: Scatter plots

* -log
* Log-log

#### Part 3: Drawing lines

* time series
* (Lambda?)
* scatter plots
* summary
* prediction


#### Part 4: smoothers?

#### Part 5: Facetings


#### Part 6: Correlations

* Summarizing a scatterplot with a single number
* Anscombes quarter
* r
* R^2?



#### Part 7: Putting it all together
scatter plot with line, R, etc


#### Part 8: Introduction to high dimensional data

* Big data vs high dimensional data
* Scatter plot matrices
* Correlation matrices

#### Part 9: Distributions

* histograms
* density plot
* comparing 2 histograms
* Amino acids
* Gene length
* Bones in mammal bodies

#### Part 10: Summarizing distributions numerically

* mean
* median
* SD
* boxplot for 1 group?

#### Part 11: Boxplots

* Sparrow telomeres stuff?

#### Part 12: Inference by eye

* Standard error and confidence interval
* Sparrow telomeres

#### Part 13: P values

* betatropin case studies


### BOOK 3: (Dimension reduction, or boiling down the numbers)

* PCA?
* Clustering 
* Euclidean distance (cover this first before genetic dist?)


### BOOK 4: Sequences

* Pairwise alignment 
* pairwise alignmen -> PID etc converts sequences to numbers

### BOOK 5: "Tree thinking"



## Ideal features of book

Some of this is available for RMarkdown but not yet part of bookdown

**Available - bookdown**
* special notices 
  + https://bookdown.org/yihui/bookdown/custom-blocks.html

**Available - rmarkdown**
* code folding
* embed webpages -> google sheets, google forms, BLAST, etc
* horizontal lines ("horizontal rule")

**Not available / unknown**
* boxed text


## Test code

```{r}
library(Stat2Data)
data(SeaIce)
```


### This is a default-sized plot

```{r, echo = F}
plot(Extent ~ Year, data = SeaIce, main = "Default")  # Note, both words capitalized.
```


### fig.height = 5, fig.width = 7

```{r, fig.height = 5, fig.width= 7, echo = F}
plot(Extent ~ Year, data = SeaIce)  
```

### fig.height = 4, fig.width = 6

```{r, fig.height = 4, fig.width= 6, echo = F}
plot(Extent ~ Year, data = SeaIce)  
```

