---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Barplots

Data from https://useast.ensembl.org/Homo_sapiens/Info/Annotation
Wed Jun 30 23:53:30 2021"

Assembly	GRCh38.p13 (Genome Reference Consortium Human Build 38), INSDC Assembly GCA_000001405.28, Dec 2013

"This site provides a data set based on the December 2013 Homo sapiens high coverage assembly GRCh38 from the Genome Reference Consortium. This assembly is used by UCSC to create their hg38 database. The data set consists of gene models built from the genewise alignments of the human proteome as well as from alignments of human cDNAs using the cDNA2genome model of exonerate."


Small non coding genes	4,865
Long non coding genes	16,896 (incl 307 readthrough)
Misc non coding genes	2,221
Pseudogenes	15,228 (incl 6 readthrough)

```{r}
number <- c(20442,4865,16896,2221,15228)
type <- c("coding","snc","lnc","misc_nc","pseudo")
genome_stats <- data.frame(type,number)
```

miniassignment - values that are out of order put into order
```{r}
ggbarplot(data = genome_stats,
          y = "number",
          x = "type", 
         fill = "type",
         order = c("coding","lnc","pseudo","snc","misc_nc"))
```

