---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Error plots


```{r}
library(compbio4all)
data(frogarms)

library(ggplot2)
library(ggpubr)
library(cowplot)
```


## Plot means with error bars

Super hand function ggerrorplot()
Default is mean +/- 1 standard error
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex")
```



Mean and se; note sd is not often used in publications
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_sd")
```


Mean and 95% confidence interval
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci")
```


Plot your data and original data
```{r}
#your data
gg.frogarms <- ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci")

#all of the data
gg.all.frogs <- ggerrorplot(data = frogarms, #change data
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci")

cowplot::plot_grid(gg.frogarms, gg.all.frogs)
```

Set colors
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci",
          color = "sex")
```


Add raw data.  kinda crazy
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci",
          color = "sex",
          shape = "sex",
          add = "point")
```


Jitter raw data.  even crazier
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci",
          color = "sex",
          add = "jitter")
```



Back to just the means Increase size
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci",
          color = "sex",
          size = 1.5) #
```



Move legend to the bottom

?set all of this stuff as eval = F and have students figur eout how to add it?
m aybe not - goal is just to do "1st encounter"
```{r}
ggerrorplot(data = frogarms,
          y = "mass",
          x = "sex",
          desc_stat = "mean_ci",
          color = "sex",
          size = 1.5,
          xlab = "Sex",
          ylab = "Mass (g)",
          legend = "bottom") #
```






## arms
```{r}
ggerrorplot(data = frogarms,
          y = "arm",
          x = "sex",
          desc_stat = "mean_ci",
          color = "sex",
          size = 1.5,
          xlab = "Sex",
          ylab = "Mass (g)",
          legend = "bottom") #
```


