---
output: html_document
editor_options: 
  chunk_output_type: console
---
#  Bookdown Notes {#bookdownnotes}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(bookdownnotes). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

## Inline R code

### Code not run, just for example

 Because of this, some lynx data is embedded within R and easy to access: all you have to do is type `data(lynx)`

## Redact expressions or lines of code

https://bookdown.org/yihui/rmarkdown-cookbook/hook-hide.html

Expressions
```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- x[!grepl('# SECRET!!$', x)]

    hook_source(x, options)
  })
})
```



Lines of code
```{r, include=FALSE}
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    #x <- x[!grepl('# SECRET!!$', x)]
    x <- xfun::split_lines(x)  # split into individual lines
    x <- x[!grepl("# SECRET!!$", x)]
    x <- paste(x, collapse = "\n")  # combine into a single string

    hook_source(x, options)
  })
})
```

## Do not stop on error

https://bookdown.org/yihui/rmarkdown-cookbook/opts-error.html

```{r, error=TRUE}
1 + "a"
```

## Style code blocks and text output

https://bookdown.org/yihui/rmarkdown-cookbook/chunk-styling.html#chunk-styling


```{r class.source="bg-danger", class.output="bg-warning"}
mtcars[1:5, "mpg"]
```


## Embed a web page

https://bookdown.org/yihui/rmarkdown-cookbook/include-url.html

## Emojii



```{r}
emo::ji("warning")
```

`r emo::ji("warning")`

https://unpkg.com/emojilib@3.0.2/dist/emoji-en-US.json 


Warning
`r emo::ji("warning")` **Note:** Unlike a spreadsheet you cannot edit the data when is called up using `View()` `r emo::ji("warning")`


Question
`r emo::ji("question")` There's a problem with my graph though - can you spot it?  It only really becomes apparent when you connect the dots with a line. `r emo::ji("question")`


## Comment out text

<!-- your comment -->

hioghlight text then Cmd + Shift + C






## Review biology through charts

chargaff1951

Higgs 2009 amino acids


## TODO

go through all my repos to look for addiationl code

blast-stats project
https://github.com/brouwern/blaststats

https://github.com/brouwern/carnivorephylo

https://github.com/brouwern/CaseStudiesEcoStats


# Packages that could be worth learning

## Formatting output stats: xplain
http://www.zuckarelli.de/xplain/index.html
http://www.zuckarelli.de/xplain/xplain_cheatsheet.pdf



## Processing images: magick
https://cran.r-project.org/web/packages/magick/vignettes/intro.html


This has a lot of dependencies and can take a long time to set up.
```{r eval = F}
install.packages("magick",dependencies = T)
```

Here's an example of how to resize some png files used in this book
```{r, eval = F}
#get names of files
files <- list.files(path = "./images/Medley1998csv",pattern = "Medley1998",full.names = T)

#make sure i just have png files, not the ppt
files <- files[grep("png",files)]

#look at file
print(image_read(files[1]))

newfiles <- gsub("Medley1998csv-","Medley1998csv-800x",files)

for(i in 1:length(files)){
  image_write(image_resize(image_read(files[i]), "800x")
, path = newfiles[i]
, format = "png")
}

```



## Two-column formatting

https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html



