---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 
TODO: move
The Montserrat Oriole only occurs on the island of Montserrat.  It therefore is an entirely closed population.  Its poplation dynamics can therefore be described by this simple eqation.

In the 1990s there were several volcanic erptions on Montserrat which reduced the Montserrat Oriole's habitat and killed off the insects it eats.  $\Delta N$ was therefore negative due to a low number of births of new birds (B near 0) and a high death rate due to startvation.



```{r}
yr.ss <- 1969:1985
n.flwr <- c(195,0,206,1,72,797,17,24,NA,284,436,NA,NA,2262,903,84,163)
precip.6mo <- c(1039,782,1376,518,388,785,453,958,
 584,393,1477,2751,542,1955,83,417,1167)

precip.12mo <- c(NA,1576,2440,774,435,1376,901,1244,733,762,2036,3147,868,2687,955,851,1461)

precip.24mo <- c(NA,NA,4016,3214,1209,1811,2277,2145,1977,1495,2798,5183,4015,3555,3642,1806,2312)

plot(n.flwr ~ yr.ss)
plot(n.flwr ~ precip.6mo)
plot(n.flwr ~ precip.12mo)
plot(n.flwr ~ precip.24mo)
```

