---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Errorplot datasets

Misc. notes on data useful for demonstrating errorplots


```{r}
library(Stat2Data)
data(Tadpoles)
```

Venesky MD, Hanlon SM, Lynch K, Parris MJ, Rohr JR. (2013) "Optimal digestion theory does not predict the effect of pathogens on intestinal plasticity," Biol Lett 9: 20130038. http://dx.doi.org/10.1098/rsbl.2013.0038

# error plot
MASS	eagles	Foraging Ecology of Bald Eagles
Knight, R. L. and Skagen, S. K. (1988) Agonistic asymmetries and the foraging ecology of Bald Eagles. Ecology 69, 1188–1194.



```{r, eval = F}
install.packages("DAAG")
install.packages("carData")
install.packages("Stat2Data")
library(DAAG)
library(carData)
data(moths) # errorplot
```

```{r}
library(DAAG)
data(moths) # errorplot
```



```{r, eval = F}

data(Depredations) # error plot
plot(number ~ longitude, data = Depredations)

```
Harper, Elizabeth K. and Paul, William J. and Mech, L. David and Weisberg, Sanford (2008), Effectiveness of Lethal, Directed Wolf-Depredation Control in Minnesota, Journal of Wildlife Management, 72, 3, 778-784. http://dx.doi.org/10.2193/2007-273



## errorplot

```{r, eval = F}


```


```{r}
library(Stat2Data)

#errorplot
data(Tadpoles)
data(Sparrows)



# percentage over time
data(SeaSlugs)
plot(SeaSlugs)



data(Pines) # includes deer damage

data(MouseBrain)  #counts


#survival
data(Leafhoppers)


data(BirdCalcium)
BirdCalcium # 


library(stevedata)
data(Amyloid)
Amyloid # errorplot
```

HawkTail
HawkTail2
Goldenrod
FruitFlies2
FruitFlies
FlightResponse

library(Stat2Data)
Cuckoo # errorplot
CrabShip # errorplot
FishEggs # errorplot


CO2 {datasets}	R Documentation
Carbon Dioxide Uptake in Grass Plants

MASS	genotype	Rat Genotype Data # errorplot?


## Health-related

boot
cd4 # paired t-test



library(Stat2Data)
DiabeticDogs # repeated measureas
Forbath, N., A. B. Kenshole, and G. Hetenyi, Jr. (1967),"Turnover lactic acid in normal and diabetic dogs calculated by two tracer methods," Am. J. Physiol. v. 212, pp.1179 - 1183.

```{r}

```

