---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Time series / climate changes

Misc. notes on data related to time series and climates change

### Stat2Data

library(Stat2Data)

SeaIce
CO2SouthPole # time series
CO2Hawaii    # time series
CO2Germany   # time series, 1 year, daily
CO2    

Could stack CO2SouthPole, CO2Hawaii, CO2Germany

Reshape SeaIce


### fpp2

"Do you want to install from sources the package which needs compilation? (Yes/no/cancel) " 
```{r, eval = F}
install.packages("fpp2") # has lots of dependencies
```


```{r}
library(fpp2)
data(maxtemp) # climate change
plot(maxtemp)
```



### Package astsa: Applied Statistical Time Series Analysis



```{r}
# Spectral envelope analysis
#install.packages("astsa")
library(astsa)
data(EBV) #Entire Epstein-Barr Virus (EBV) Nucleotide Sequence
```


blood	Daily Blood Work
bnrf1ebv	Nucleotide sequence - BNRF1 Epstein-Barr
bnrf1hvs	Nucleotide sequence - BNRF1 of Herpesvirus saimiri

* cardox	ts objec;  Monthly Carbon Dioxide Levels at Mauna Loa
  
cmort	Cardiovascular Mortality from the LA Pollution study
dna2vector	Convert DNA Sequence to Indicator Vectors
EBV	Entire Epstein-Barr Virus (EBV) Nucleotide Sequence
flu	ts object; Monthly pneumonia and influenza deaths in the U.S., 1968 to...
fmri	fMRI - complete data set
fmri1	fMRI Data Used in Chapter 1
globtemp	Global mean land-ocean temperature deviations to 2015
globtempl	Global mean land (only) temperature deviations to 2015
gtemp	Global mean land-ocean temperature deviations
gtemp2	Global Mean Surface Air Temperature Deviations
gtemp_land	Global mean land temperature deviations - updated to 2017
gtemp_ocean	Global mean ocean temperature deviations - updated to 2017
Hare	Snowshoe Hare
HCT	Hematocrit Levels
hor	Hawaiian occupancy rates
lap	LA Pollution-Mortality Study

Lynx	Canadian Lynx
http://people.whitman.edu/~hundledr/courses/M250F03/M250.html
http://people.whitman.edu/~hundledr/courses/M250F03/LynxHare.txt
NOTE about the data: I have not been able to verify this data, but this is the data (or rather the graph) that is always cited. This particular set of data came from scanning in the graph from Odum's "Fundamentals of Ecology", p. 191 which is often cited. Odum says that his graph is taken from MacLulich's "Fluctuations in the numbers of varying hare", 1937, which is not widely available. Some authors caution that this data is actually a composition of several time series, and should probably not be analyzed as a whole, and that some of the lynx data was actually missing. It is said that the data was collected from Hudson's Bay historical records, and does not reflect animal populations, but rather the number of pelts turned in for trading (a large number of which came from Native Americans- mentioned because there were some medical outbreaks during these years which could account for skewed data). The data is presented here with these cautions.

part	Particulate levels from the LA pollution study
PLT	Platelet Levels
polio	Poliomyelitis cases in US
rec	Recruitment (number of new fish index)
salt	Salt Profiles
saltemp	Temperature Profiles
so2	SO2 levels from the LA pollution study
soi	Southern Oscillation Index
soiltemp	Spatial Grid of Surface Soil Temperatures
sunspotz	Biannual Sunspot Numbers
tempr	Temperatures from the LA pollution study
WBC	White Blood Cell Levels
