# Multivariate 

Misc. notes on data related to multivariate analyses / visualization

MASS	waders	Counts of Waders at 15 Sites in South Africa
R.W. Summers, L.G. Underhill, D.J. Pearson and D.A. Scott (1987) Wader migration systems in south and eastern Africa and western Asia. Wader Study Group Bulletin 49 Supplement, 15–34.


MASS	crabs	Morphological Measurements on Leptograpsus Crabs
Campbell, N.A. and Mahon, R.J. (1974) A multivariate study of variation in two species of rock crab of genus Leptograpsus. Australian Journal of Zoology 22, 417–425.


```{r, eval = F}
install.packages("HistData")
library(HistData) 
```





```{r, eval = F}
# install.packages("robustbase")
library(robustbase)
data(possumDiv) # multiple regression, multivar

plot(Diversity ~ Habitat, data = possumDiv)
```

Lindenmayer, D. B., Cunningham, R. B., Tanton, M. T., Nix, H. A. and Smith, A. P. (1991) The conservation of arboreal marsupials in the montane ash forests of the central highlands of victoria, south-east australia: III. The habitat requirements of leadbeater's possum gymnobelideus leadbeateri and models of the diversity and abundance of arboreal marsupials. Biological Conservation 56, 295–315.

Lindenmayer, D. B., Cunningham, R. B., Tanton, M. T., Smith, A. P. and Nix, H. A. (1990) The conservation of arboreal marsupials in the montane ash forests of the victoria, south-east australia, I. Factors influencing the occupancy of trees with hollows, Biological Conservation 54, 111–131.
