---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Datasets - regression

Misc. notes on data related to regression.

## Repated measures
MASS	Sitka	Growth Curves for Sitka Spruce Trees in 1988
```{r}
library(MASS)
data(Sitka)
plot(size ~ Time, data = Sitka)
```

MASS	Sitka89	Growth Curves for Sitka Spruce Trees in 1989


MASS  cats	Anatomical Data from Domestic Cats # regression
	
	
	
	
## Multiple regression 

```{r}
library(Stat2Data)
data(MetabolicRate)

plot(Mrate ~ BodySize, data = MetabolicRate)
plot(CO2ppm ~ BodySize, data = MetabolicRate)
plot(CO2ppm ~ Mrate, data = MetabolicRate)
```



```{r, eval = F}
library(Stat2Data)
Fertility ## regression, multiple regression
```


```{r}
library(Stat2Data)
data(ElephantsMF)
data(ElephantsFB)
ElephantsMF # regression, multiple regression
ElephantsFB # regression, multiple regression

plot(Height ~ Age, data = ElephantsFB)
```


Caterpillars # regression, multiple regression; nonlinear?
BirdNest # regression, multiple regression

## GLM


```{r, eval = F}
# install.packages("robustbase")
library(robustbase)
data(possumDiv) # multiple regression, multivar

plot(Diversity ~ Habitat, data = possumDiv)
```

Lindenmayer, D. B., Cunningham, R. B., Tanton, M. T., Nix, H. A. and Smith, A. P. (1991) The conservation of arboreal marsupials in the montane ash forests of the central highlands of victoria, south-east australia: III. The habitat requirements of leadbeater's possum gymnobelideus leadbeateri and models of the diversity and abundance of arboreal marsupials. Biological Conservation 56, 295–315.

Lindenmayer, D. B., Cunningham, R. B., Tanton, M. T., Smith, A. P. and Nix, H. A. (1990) The conservation of arboreal marsupials in the montane ash forests of the victoria, south-east australia, I. Factors influencing the occupancy of trees with hollows, Biological Conservation 54, 111–131.

```{r}

```

