---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Datasets

Misc. notes on data that may be interesting

```{r}
library(robustbase)
data(vaso) #Vaso Constriction Skin Data Set

```


multivariate; gene expression, clustering
https://archive.ics.uci.edu/ml/datasets/Mice+Protein+Expression


multivar, regression
```{r}
library(robustbase)
data(toxicity)
```

Maguna, F.P., Núñez, M.B., Okulik, N.B. and Castro, E.A. (2003) Improved QSAR analysis of the toxicity of aliphatic carboxylic acids; Russian Journal of General Chemistry 73, 1792–1798.

library(robustbase)
data(biomassTill)

library(robustbase)
data(possumDiv)

library(robustbase)
phosphor
This dataset investigates the effect from inorganic and organic Phosphorus in the soil upon the phosphorus content of the corn grown in this soil, from Prescott (1975).

library(robustbase)
NOxEmissions
A typical medium sized environmental data set with hourly measurements of NOx pollution content in the ambient air.




```{r}
library(robustbase)
data(salinity)
```
This is a data set consisting of measurements of water salinity (i.e., its salt concentration) and river discharge taken in North Carolina's Pamlico Sound, recording some bi-weekly averages in March, April, and May from 1972 to 1977. This dataset was listed by Ruppert and Carroll (1980). In Carrol and Ruppert (1985) the physical background of the data is described.





Fitch

```{r}
library(Stat2Data)
data(Fitch)
plot(Palate ~ Skull, data = Fitch) # linear
plot(Skull ~ Wt, data = Fitch)     # non-linear
```

Fitch, W. Tecumseh (2000), "Skull dimensions in relation to body size in nonhuman mammals: The causal bases for acoustic allometry," Zoology, 103, 40-58.


## Non-linear curve

```{r}
library(Stat2Data)
data(Fluorescence)
plot(ProteinProp ~ Calcium, data = Fluorescence)
```



## Noisy correlation

```{r}
library(Stat2Data)
data(MothEggs)
plot(Eggs ~ BodyMass, data = MothEggs)
```



## Regression; quantile regression ; climate change

```{r}
library(Stat2Data)
data(LeafWidth)
plot(LWRatio ~ Year, data = LeafWidth)
```

Guerin, G., Wen, H., Lowe, A. (2012), "Leaf morphology shift linked to climate change," Biol. Lett., 8, doi: 10.1098/rsbl.2012.0458




## Logistic regression
```{r}
library(Stat2Data)
data(Gunnels)
```




## Regression; Fitness
```{r}
library(Stat2Data)
data(GlowWorms)
plot(GlowWorms)
```

Hopkins J, Baudry G, Candolin U, Kaitala A. (2015), "I'm sexy and I glow it: female ornamentation in a nocturnal capital breeder," Biol. Lett. 11: 20150599.
http://dx.doi.org/10.1098/rsbl.2015.0599












BrainpH #

# 2 x 2 table
TMS # TMS Effects of transcranial magnetic stimulation (TMS) on migraine headaches
Based on results in R. B. Lipton, et. al. (2010) “Single-pulse Transcranial Magnetic Stimulation for Acute Treatment of Migraine with Aura: A Randomised, Double-blind, Parallel-group, Sham-controlled Trial," 9(4):373-380.



library(Stat2Data)
CancerSurvival




      






ButterfliesBc # regression, climate chnage
Digitized data from plots in Bowden, J. et al., "High-Arctic butterflies become smaller with rising temperatures", published in Biology Letters 11: 20150574
```{r}
library(Stat2Data)
data(ButterfliesBc)
plot(Temp ~ Wing, data = ButterfliesBc)
```




```{r, eval = F}
install.packages("stevedata")
library(stevedata)
data(Datasaurus)
Datasaurus
```





```{r, eval = F}
install.packages("survival")
```


```{r}
library(survival)
```


```{r, eval = F}
install.packages("MASS")
```







MASS	caith	Colours of Eyes and Hair of People in Caithness


MASS	GAGurine	Level of GAG in Urine of Children



MASS	snails	Snail Mortality Dat


HistData	Arbuthnot	Arbuthnot's data on male and female birth ratios in London from 1629-1710.

HistData	ChestSizes	Chest measurements of 5738 Scottish Militiamen
HistData	Cholera	William Farr's Data on Cholera in London, 1849




## Heritablity 

HistData	Galton	Galton's data on the heights of parents and their children
HistData	GaltonFamilies	Galton's data on the heights of parents and their children, by child
HistData	PearsonLee	Pearson and Lee's data on the heights of parents and children classified by gender




## Mendelian genetics 

hwde	IndianIrish	Observed genotype frequencies at MN and S loci, for 2 populations
hwde	mendelABC	Mendel's F2 trifactorial data for seed shape (A: round or wrinkled), cotyledon color (B: albumen yellow or green), and seed coat color (C: grey-brown or white)






## Genomics
ade4 bacteria
AA, codon, nucleotide counts for 43 spp of bacteria




## Time series data library

https://pkg.yangzhuoranyang.com/tsdl/reference/tsdl.html

```{r}

```



## Data sets 

ChickWeight   Weight versus age of chicks on different diets
chickwts 	Chicken Weights by Feed Type
CO2 	Carbon Dioxide Uptake in Grass Plants
co2 	Mauna Loa Atmospheric CO2 Concentration
InsectSprays   Effectiveness of Insect Sprays
iris 	Edgar Anderson's Iris Data
iris3 	Edgar Anderson's Iris Data
islands 	Areas of the World's Major Landmasses
LakeHuron   Level of Lake Huron 1875-1972
Loblolly   Growth of Loblolly pine trees
lynx 	Annual Canadian Lynx trappings 1821-1934
npk   Classical N, P, K Factorial Experiment
Orange   Growth of Orange Trees
OrchardSprays 	Potency of Orchard Sprays
PlantGrowth   Results from an Experiment on Plant Growth
precip 	Annual Precipitation in US Cities
rivers   Lengths of Major North American Rivers
rock 	Measurements on Petroleum Rock Samples
trees   Girth, Height and Volume for Black Cherry Trees
women   Average Heights and Weights for American Women


http://www.zoology.ubc.ca/~whitlock/ABD/teaching/datasets.html

